Exploiting the Social Graph: Increasing Engagement in Collaborative Interactive Evolution

### 2017-01-27 2017-CEC-Graph-Based-Population-For- IEC/paper.tex
Deleting some addressed issues
And conclusion is still missing, so I will wait until it is ready to
work on it. Ping me later when you do it. Maybe start working on the
next paper, too, right?
Comments addressed, going to Figures and Conclusion.
@jj most comments addressed and a few question, please have a look.

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users. % Adding this connecting
                                % to paper's objetives - JJ
                                % Thanks yes -Mario

@@

\item {\bf Collection} This type of node represents the collections belonging
to users, their attributes are {\tt id}, 
{\tt name}, {\tt created} and {\tt is\_private} indicating if contents
belonging to this collection can be shared with others. Collections can contain
phenotypes and a single phenotype could be shared by many collections. % This is why they are not attributes - Mario 
\end{itemize}

% Add something at the end linking with the next paragraph
% Done - Mario
The interaction between these entities and how they are represented by relationships 
between nodes are presented in the next subsection.

@@

\begin{itemize}
\item {\bf Likes} This relation describes the interaction between a user and
a phenotype in which a rating value is assigned this is represented by 
the {\tt rate} attribute. In the current application a user is allowed to rate
a phenotype multiple times, so the date of the interactions is also of
interest.
%this is a bipartite graph them. - JJ
% Do you mean like a figure?  - Mario

\item {\bf Knows} The relation connects two users that know each
other. 
% This relation is "unique" meaning that it cannot be repeated
% between a pair of nodes.  % Not too clear what you mean by that - JJ
% I mean is not a multigraph or you can not have two equal relationships between nodes. - Mario
 % Watching the graph structure of Neo4J, I can now see that is actually a directed graph,
 % There is even cases where knows relationship only goes in one direction, this could be because
 % a user does not publishes their friends or one only follows another but are not friends of each other.  
% Deleting the restriction, I think is not important and can be seen as a graph.

@@

\item {\bf Shared} When a user takes an phenotype from a friend's collection, this interactions 
is described by a {\tt share} relationship between the user and the
phenotype.  % Still think that user and individual is confusing - JJ
            % Agreed corrected -Mario
            % I was thinking about calling them Solutions or Candidate Solutions,

@@

\subsubsection{ Base deployment}
After a week of the announcement the total number of volunteers for this experiment was 53 registered users. 
The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
experiment a total of 500 new phenotypes where generated. The total number of evaluated
phenotypes for each volunteer is presented in figure
\ref{fig:top-ranked-participation} with a blue line. % Ranked by... - JJ

\subsubsection{ Non Graph Gamification deployment}
The total number of volunteers for this experiment was very close to those obtained in the previous
experiment: 54 registered users. The total number of nodes generated in the graph was 648, 
with a total of 2596 relationships. These numbers are close to those obtained in the base experiment,
with a marginal increase in participation. The increased participation also helped in the creation of
more phenotypes for a total of 648. The total number of evaluated
phenotypes for each volunteer is presented in figure \ref{fig:top-ranked-participation} with a red line.

@@

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/user_known_2.png}
    \caption{User network view of the Base deployment}
    \label{fig:B-network}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/user_known_3.png}
    \caption{User network view of the Graph Gamification deployment}
    \label{fig:GG-network}
\end{figure}
\begin{figure*}[!t]
    \centering
        \includegraphics[width=5.5in]{img/comparison.png}
    \caption{Top ranked number phenotypes rated by each user using a
     logarithmic scale.} % Explain a little more and label axes. The
                         % three lines have the same label,
                         % "participation". Not too clear - JJ
    \label{fig:top-ranked-participation}
\end{figure*}

When monitoring the experiments, using the graph database and the Cypher query language 
provided interesting views on the connected data difficult to replicate with a relational database. 
As an example we can consider comparing the difference between the number of volunteers in deployments
B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively) additional
metrics and patterns are available when they are treated as a
network. %Please read the whole sentence and maybe rewrite - JJ
The same can be applied to 
the population, or the interaction as a whole. The graph could be used also by the evolutionary
algorithm, for instance when creating a new phenotype 
parents could be selected only if there is a minimum distance between them in the graph.

@@

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\bibliographystyle{IEEEtran}
\bibliography{evospace-i,volunteer,biblio,geneura}

>>>>

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users.

@@

\item {\bf Collection} This type of node represents the collections belonging
to users, their attributes are {\tt id}, 
{\tt name}, {\tt created} and {\tt is\_private} indicating if contents
belonging to this collection can be shared with others. Collections can contain
phenotypes and a single phenotype could be shared by many
collections. % This is why they are not attributes - Mario
% You can simply have an attribute that is a list of the collections
% it belongs to - JJ
\end{itemize}

The interaction between these entities and how they are represented by relationships 
between nodes are presented in the next subsection.

@@

\begin{itemize}
\item {\bf Likes} This relation describes the interaction between a user and
a phenotype in which a rating value is assigned this is represented by 
the {\tt rate} attribute. In the current application a user is allowed to rate
a phenotype multiple times, so the date of the interactions is also of
interest.
%this is a bipartite graph them. - JJ
% Do you mean like a figure?  - Mario
% No, it's a kind of graph - JJ

\item {\bf Knows} The relation connects two users that know each
other. 
% This relation is "unique" meaning that it cannot be repeated
% between a pair of nodes.  % Not too clear what you mean by that - JJ
% I mean is not a multigraph or you can not have two equal relationships between nodes. - Mario
 % Watching the graph structure of Neo4J, I can now see that is actually a directed graph,
 % There is even cases where knows relationship only goes in one direction, this could be because
 % a user does not publishes their friends or one only follows another but are not friends of each other.  
% Deleting the restriction, I think is not important and can be seen
% as a graph.
% It's a directed, bipartite graph then - JJ

@@

\item {\bf Shared} When a user takes an phenotype from a friend's collection, this interactions 
is described by a {\tt share} relationship between the user and the
phenotype.  % Still think that user and individual is confusing - JJ
            % Agreed corrected -Mario
            % I was thinking about calling them Solutions or Candidate
            % Solutions,
% Solutions is OK, but also phenotype - JJ

@@

\subsubsection{ Base deployment}
After a week of the announcement the total number of volunteers for this experiment was 53 registered users. 
The total number of nodes generated in the graph was 595, with a total of 2220 relationships. In this
experiment a total of 500 new phenotypes where generated. The total number of evaluated
phenotypes for each volunteer is presented in figure
\ref{fig:top-ranked-participation} in blue. Users are ranked by
the number of phenotypes they rated, in the \emph{x} axis is the rank, and in the \emph{y} axis 
the number of phenotypes evaluated using a logarithmic scale.   % Ranked by... - JJ

\subsubsection{ Non Graph Gamification deployment}
The total number of volunteers for this experiment was very close to those obtained in the previous
experiment: 54 registered users. The total number of nodes generated in the graph was 648, 
with a total of 2596 relationships. These numbers are close to those obtained in the base experiment,
with a marginal increase in participation. The increased participation also helped in the creation of
more phenotypes for a total of 648. The total number of evaluated
phenotypes for each volunteer is presented in figure \ref{fig:top-ranked-participation} in red.

@@

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/user_known_2.png}
    \caption{User network view of the Base deployment}
    \label{fig:B-network}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/user_known_3.png}
    \caption{User network view of the Graph Gamification deployment}
    \label{fig:GG-network}
\end{figure}
\begin{figure*}[!t]
    \centering
        \includegraphics[width=5.5in]{img/comparison.png}
    \caption{Users are ranked by
		the number of phenotypes they rated, in the \emph{x} axis is the rank, and in the \emph{y} axis 
		the number of phenotypes evaluated using a logarithmic scale. } % Explain a little more and label axes. The
                         % three lines have the same label,
                         % "participation". Not too clear - JJ
                         % TO DO -Mario
    \label{fig:top-ranked-participation}
\end{figure*}

When monitoring the experiments by querying the graph database,  the Cypher query language 
provided interesting views on the data that it would have been difficult to replicate with a relational database.
For instance, to compare the number of volunteers in deployments
B and GG (shown in figures \ref{fig:B-network} and \ref{fig:GG-network} respectively)  a simple query
provided a view of the social network of volunteers. 
%Please read the whole sentence and maybe rewrite - JJ
% Done - Mario

The same can be applied to 
the population, or the interaction as a whole. The graph could be used also by the evolutionary
algorithm, for instance when creating a new phenotype 
parents could be selected only if there is a minimum distance between them in the graph.

@@

\bibliographystyle{IEEEtran}
\bibliography{evospace-i,volunteer,biblio,geneura}
